Accelerated evolution of convective simulations

### 2018-06-14 tex/bvp_paper.tex
almost ready for resubmission

Unfortunately, astrophysical convective systems are stiff in more than one 
timescale. Specifically, the Peclet number (Pe), the ratio of the thermal
diffusion timescale to the convective velocity timescale, is large.
In a high Pe system, many convective dynamical timescales must pass before 
the domain reaches thermal equilibrium, a steady state in which
evolution of the thermal structure of the convective region has ceased.
Fully implicit methods cannot be used to address the stiffness resulting from
the discrepancy between convective and thermal timescales; convective flows are
highly nonlinear, and implicit methods are bound by the nonlinear flow timescales
\cite{viallet&all2011, viallet&all2013, viallet&all2016}.

Resolving dynamics in atmospheres which are sufficiently
thermally equilibrated remains a challenging problem.
Solar convection is a prime example of this phenomenon, as
dynamical timescales in the solar convective zone are relatively short 
(convection overturns every $\sim 5$ minutes at the solar surface)
compared to the Sun's thermal evolution timescale, which is O($10^7$) years
\cite{stix2003}.  
In such a system, it is impossible to resolve the convective dynamics while also
meaningfully evolving the thermal structure of the system using
traditional timestepping techniques alone.
As modern simulations aim to model natural, high-Pe convection
in the high-Rayleigh-number (Ra) regime,
the thermal diffusion timescale ($t_{\kappa}$, defined in section \ref{sec:ae}) 
becomes intractably large compared to dynamical timescales 
such as the freefall time ($t_{\text{ff}}$, defined in section \ref{sec:experiment})
\cite{anders&brown2017}, 
\begin{equation}
\frac{t_{\kappa}}{t_{\text{ff}}} \propto (\text{Ra})^{1/2}.
\end{equation}
Furthermore, as dynamical and thermal timescales separate, 
simulations become more turbulent. Capturing appropriately resolved
turbulent motions requires finer grid meshes and smaller timesteps.
Thus, the progression of simulations into the high-Ra
regime of natural convection is slowed by two simultaneous effects: timestepping
through a single convective overturn time becomes more computationally expensive
and the number of overturn times required for systems to reach thermal equilibration
grows.

@@

In this work, we study a method of achieving accelerated evolution of
convective simulations. We couple measurements of the dynamics of unequilibrated
convective simulations with knowledge about energy balances in the desired solution
to self-consistently adjust the mean vertical thermodynamic profile towards its evolved state. 
While a technique of this kind has been used previously \cite{hurlburt&all1986},
the details of implementation, the convergence properties, and whether or not the
accelerated evolution state corresponds to a standard evolution state are not documented.
In section \ref{sec:experiment}, we describe our convective simulations and
numerical methods. In section \ref{sec:ae}, we describe the procedure for 
achieving accelerated evolution. In
section \ref{sec:results}, we compare accelerated evolution solutions
to solutions obtained from standard evolution through a full thermal diffusion timescale. 
In section \ref{sec:speedups}, we compare the numerical cost of accelerated
solutions to solutions that are evolved through a thermal diffusion timescale.
Finally,
in section \ref{sec:extensions}, we offer concluding remarks and
discuss extensions of the methods presented here.

@@

\section{Experiment}
\label{sec:experiment}
In this work we study a simple form of thermal convection:
incompressible \RB convection under the Oberbeck-Boussinesq approximation,
such that the fluid
has a constant kinematic viscosity ($\nu$), thermal diffusivity ($\kappa$), and coefficient
of thermal expansion ($\alpha$). The density of the fluid is a constant, $\rho_0$,
except in the buoyancy term, where it is $\rho = \rho_0(1  - \alpha T_1)$.
The gravitational acceleration, $\bm{g} = - g\bm{\hat{z}}$, is constant.
The equations of motion are \cite{spiegel&veronis1960}:
\begin{gather}
\DivU = 0, 
	\label{eqn:incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \bm{u}\cdot\grad\bm{u} =
-\frac{1}{\rho_0}\grad P - g( 1 - \alpha T_1)\hat{z} + \nu\grad^2\bm{u}, 
	\label{eqn:dim_bouss_momentum}
\\
\frac{\partial T_1}{\partial t} + \bm{u}\cdot\grad(T_0 + T_1) = \kappa\grad^2 T_1,
	\label{eqn:dim_bouss_energy}
\end{gather}
where $\bm{u} = u\bm{\hat{x}} + v\bm{\hat{y}} + w\bm{\hat{z}}$ is the velocity, 
$T = T_0(z) + T_1(x, y, z, t)$ are the initial and fluctuating components of temperature, 
and $P$ is the kinematic pressure. The initial temperature profile decreases
linearly with height.
We non-dimensionalize these equations such that
length is in units of the layer height ($L_z$),
temperature is in units of the initial temperature jump across the layer ($\Delta T_0 = L_z \partial_z T_0$), 
and velocity is in units of the freefall velocity ($v_{\text{ff}} = \sqrt{\alpha g L_z^2 \partial_z T_0}$).
By these choices, one time unit is a freefall time ($t_{\text{ff}} = L_z/v_{\text{ff}}$).
We introduce a reduced kinematic pressure,
$\varpi \equiv (P / \rho_0 + \phi + |\bm{u}|^2 / 2) / v_{\text{ff}}^2$, where the gravitational
potential, $\phi$, is defined such that $\bm{g} = -\grad \phi$. 
In non-dimensional form, and substituting 
$\bm{u}\cdot\grad\bm{u} = \grad(|\bm{u}|^2/2) - \bm{u}\times(\grad\times\bm{u})$
and $\grad^2\bm{u} = -\grad\times(\grad\times\bm{u})$, Eqns.~\ref{eqn:dim_bouss_momentum} \& \ref{eqn:dim_bouss_energy}
become
\begin{align}
\frac{\partial \bm{u}}{\partial t} + \grad \varpi - T_1\hat{z} + \mathcal{R}\grad\times\bm{\omega} &= \bm{u}\times\bm{\omega},
	\label{eqn:bouss_momentum}
\\
\frac{\partial T_1}{\partial t} - \mathcal{P}\grad^2 T_1 + w \frac{\partial T_0}{\partial z} &= - \bm{u}\cdot\grad T_1,
	\label{eqn:bouss_energy}
\end{align}
where $\bm{\omega} = \grad \times \bm{u}$ is the vorticity.
The dimensionless control parameters $\mathcal{R}$ and $\mathcal{P}$ 
are set by the Rayleigh (Ra) and Prandtl (Pr) numbers,
\begin{equation}
\mathcal{R} \equiv \sqrt{\frac{\text{Pr}}{\text{Ra}}}, \qquad \mathcal{P} \equiv \frac{1}{\sqrt{\text{Pr}\,\text{Ra}}}, \qquad
\text{Ra} = \frac{g \alpha L_z^4 \partial_z T_0}{\nu\kappa} = \frac{(L_z\,v_{\text{ff}})^2}{\nu\kappa}, \qquad \text{Pr} = \frac{\nu}{\kappa}.
\end{equation}
We hold Pr $= 1$ constant throughout this work, such that $\mathcal{P} = \mathcal{R}$.
$\mathcal{P}$ and $\mathcal{R}$ are related to the inverse Reynolds and Peclet numbers of the
evolved flows.

In Eqns.~(\ref{eqn:incompressible}), (\ref{eqn:bouss_momentum}), \& (\ref{eqn:bouss_energy}),
linear terms are grouped on the left-hand side of the equations, while nonlinear terms
are found on the right-hand side. We timestep linear terms implicitly, and nonlinear
terms explicitly.
We utilize the 
Dedalus\footnote{\url{http://dedalus-project.org/}} 
pseudospectral framework \cite{burns&all2016} to evolve  
Eqns.~(\ref{eqn:incompressible}), (\ref{eqn:bouss_momentum}), \& (\ref{eqn:bouss_energy}) 
forward in time
using an implicit-explicit (IMEX), third-order, four-stage 
Runge-Kutta timestepping scheme RK443 \cite{ascher&all1997}.

@@

The horizontally averaged profiles of the vertical conductive flux, 
F$_{\kappa} = \angles{-\kappa\grad(T_0 + T_1)}_{x,y}$, and the vertical convective enthalpy flux,
F$_{\text{E}} = \angles{w(T_0 + T_1)}_{x,y}$, are the basis of the AE method.
Here we use $\angles{}_{x,y}$ to represent a horizontal average. We measure
both of these fluxes early in a simulation, retrieving profiles such as
those shown in Fig.~\ref{fig:time_trace}b.
As the atmosphere evolves towards
the isothermal profile specified by the upper (cold) boundary condition, excess
temperature throughout the atmosphere must leave the domain. This excess thermal
energy leaves through the upper boundary, as seen in 
Fig.~\ref{fig:time_trace}b, where the amount of flux exiting at the top of the domain
is nearly 20 times larger the flux entering the bottom of the domain. Once the atmospheric
temperature profile reaches its evolved state, the flux entering the bottom boundary
is equal to the flux exiting through the upper boundary.  In general, this 
evolution is slow in SE (Fig.~\ref{fig:time_trace}a), but AE (Fig.~\ref{fig:time_trace}c)
can rapidly advance a system whose fluxes are in a strongly disequilibrium state (Fig.~\ref{fig:time_trace}b),
into a near-equilibrium state, as shown in Fig.~\ref{fig:time_trace}d. In
this final state both boundaries conduct the same amount of flux.
The converged fluxes achieved through AE is at most 5\% different from the SE solution, 
as shown in Fig.~\ref{fig:time_trace}e. 
This is a very small difference considering the short timescales on which convergence
is reached and the strongly disequilibrium state used to inform the AE process.

@@

To properly reduce the conductive flux, we examine the 
time-stationary state, horizontally- and time-averaged
Eqns.~(\ref{eqn:bouss_momentum}) and (\ref{eqn:bouss_energy}); after 
neglecting terms which vanish due to symmetry, these equations become
\vspace{-0.2cm}
\begin{gather}
\frac{\partial}{\partial z}\angles{\varpi}_{x,y} - \angles{T_1}_{x,y}\hat{z} = \angles{\bm{u}\times\bm{\omega}}_{x,y, \text{ ev}},
	\label{eqn:bouss_BVP_momentum}
\\
\frac{\partial}{\partial z}\text{F}_{\text{E, ev}} - \mathcal{P}\frac{\partial^2}{\partial z^2} \angles{T_1}_{x,y} = 0.
	\label{eqn:bouss_BVP_energy}
\end{gather}
Here, we construct  $\text{F}_{\text{E, ev}} = \xi \text{F}_{\text{E}}$ 
and $\angles{\bm{u}\times\bm{\omega}}_{x, y,\text{ ev}} =
\xi \angles{\bm{u}\times\bm{\omega}}_{x, y}$, where we scale 
$\angles{\bm{u}\times\bm{\omega}}_{x, y} \propto \bm{u}^2$, by $\xi$ because we
scaled $\bm{u}$ by $\sqrt{\xi}$ in order to reduce the convective enthalpy flux.
With these profiles, we solve for $\angles{\varpi}_{x,y}$ and $\angles{T_1}_{x,y}$.
Convective flows
are perturbations around a thermal profile defined by these equations in the proper evolved, 
statistically stationary state, so the proper specification of
F$_{\text{conv, ev}}$ and $\angles{\bm{u}\times\bm{\omega}}_{x,y\text{, ev}}$,
fully specifies the mean thermodynamic structure of the system. 
Our choice of fixing the temperature at the top boundary and the flux at the bottom
boundary ensures that there is a unique solution for $\angles{T_1}$ given the evolved
fluxes. If flux were fixed at both boundaries, there would be infinitely many
temperature solutions. If temperature were fixed at both boundaries, 
$\text{F}_{\text{B}}$ would not be precisely known \emph{a priori}.

Thus, the AE method occurs in the following steps. First, we pause a convective simulation and
construct $\xi(z)$ from measured flux profiles in the convective domain.
We solve a 1D boundary value problem (BVP) consisting of
Eqns.~(\ref{eqn:bouss_BVP_momentum}) \& (\ref{eqn:bouss_BVP_energy})
to obtain an evolved thermodynamic profile, and its corresponding conductive flux.
As the AE BVP solve only adjusts $\angles{T}$ and $\angles{\varpi}$, we must
also appropriately adjust the temperature perturbations around the mean and the
convective velocity flows.
After adjusing the fields of a simulation in this manner, we continue timestepping forward.
For specifics on the precise implementation of the AE method, we refer
the reader to appendix \ref{appendix:recipe}.

@@

\begin{figure}[b]
\includegraphics[width=\textwidth]{./figs/nu_v_time.png}
\caption{The time evolution of Nu (black) and $\angles{T} - T_{\text{top}}$ (blue)
are shown for simulations at $S = 10^5$ in SE (a) and AE (b).
A moving average of Nu, using a centered boxcar with a span of 50 freefall time
units, is overplotted as a thick, dashed line. 
We see that the mean value of Nu reaches its equilibrium value quickly
compared to $\angles{T}$.
Regardless, fluctuations of Nu about the mean
value are much smaller in the thermally equilibrated state.
The times at which AE solves occur are denoted by arrows in (b),
and the grey boxed region is examined in more detail in Fig. \ref{fig:oscillating_plumes}.
\label{fig:nu_v_time} }
\end{figure}

@@

The Nusselt number (Nu) quantifies the efficiency of convective heat transport
and is defined as
\begin{equation}
\text{Nu} = \frac{\angles{F_{\text{conv}} + F_{\text{cond}}}}{\angles{F_{\text{cond, ref}}}}
 = \frac{\angles{wT - \mathcal{P}\partial_z T}}{\angles{- \mathcal{P} \partial_z T}},
\end{equation}
where the volume average of a quantity, $\eta$, is shown as $\angles{\eta}$.
The time evolution of Nu in AE and SE is compared to the mean temperature evolution
in Fig.~\ref{fig:nu_v_time}. In Fig.~\ref{fig:nu_v_time}a, we show the evolution
of the SE run at $S = 10^5$. The black trace shows the instantaneous value of
Nu, and the overplotted thick dashed line shows a moving time average of Nu. The
time average is taken using a center boxcar window whose width is 50 $t_{\text{ff}}$.
The mean value of Nu evolves towards its final value (shown in the horizontal
dashed black line) rapidly compared to the mean system temperature (shown in
blue), which approaches its evolved state (horizontal dashed blue line) slowly.
As such, it is possible to measure the statistically stationary mean value of
Nu after a few hundred freefall times, 
as has been done previously \cite{stevens&all2010}. However, the fluctuations
around the mean value are larger at early times when the system is not thermally
equilibrated. In Fig.~\ref{fig:nu_v_time}b, we show the AE traces of Nu and mean
temperature at $S = 10^5$. The times at which AE solves are completed are marked
by arrows. Nu reaches its evolved mean value slightly more rapidly than in the SE case,
but the early, large fluctuations away from the mean that we see in the SE
case are not observed after the AE solves.

When $S < 10^{3.67}$ in 2D and for all runs in 3D, 
the evolved system is characterized by a time-stationary value of Nu, and is thus
in a state of constant convective heat transport.
At larger $S$ in 2D, the value of Nu varies significantly over time even in the
equilibrated state (as seen in Fig.~\ref{fig:nu_v_time}). We examine the shaded
region of Fig.~\ref{fig:nu_v_time}b in more detail in the top left
panel of Fig.~\ref{fig:oscillating_plumes}, as well as a comparable time trace
at $S = 10^7$ (bottom, left). We find that these systems exhibit
large Nu during
states in which temperature fluctuations travel in their natural buoyant
direction (Fig.~\ref{fig:oscillating_plumes}, Ia \& IIa, where cold elements fall and hot elements rise).
However, when wrongly-signed temperature perturbations are entrained in an upflow or downflow
with oppositely signed fluid, Nu is suppressed (Fig.~\ref{fig:oscillating_plumes}, Ib and IIb, 
where warm fluid is pulled down by the downflow
lane, and cool fluid is drawn up by the upflow lane).
The plumes in these
systems naturally oscillate horizontally over time, switching between transport being dominated
by a counterclockwise cell, as pictured in Fig.~\ref{fig:oscillating_plumes}, and
a clockwise cell. Our choice of no-slip
boundary conditions prevents the fluid from entering a full domain shearing state 
\cite{goluskin&all2014}, and the
oscillatory nature of the plumes is stable. However, thanks to our choice of periodic
boundary conditions and despite the no-slip conditions, the full system of the
upflow and downflow plumes is free to slowly migrate to the left or right over time,
and we observe this phenomenon in our solutions.
The 2D SE simulations exhibit the same horizontally
oscillatory behavior as the AE solutions for the same initial conditions. 
This time-dependent behavior of Nu is not seen strongly in our 3D solutions,
however most 3D simulations we conducted were at low $S$ compared to the runs in
which this behavior was observed in 2D.

@@

\begin{figure}[t]
\includegraphics[width=\textwidth]{./figs/temp_comparison.png}
\caption{Comparisons of the evolved thermodynamic states of an AE and SE run
at $S = 10^{5}$ are shown.  (a) Evolved horizontally- and time-averaged 
temperature profiles, as a function of height.
(b) Probability Distribution Functions (PDFs) and their integrated
Cumulative Distribution Functions (CDFs)
of point-by-point measurements of the temperature field.
(c) The percentage difference between the mean temperature profiles as a function of height.
The difference between the mean profiles is very small, O(0.5\%).
(d) $\KS{T}$, as defined in Eqn.~\ref{eqnks_profile}, is shown. The small
difference in the mean interior temperature between AE and SE
results in a large difference between the two temperature distributions near the values
of the temperature maxima.  The spread of temperature around the maxima, which includes the 
fluctuations that drive convection, are nearly identical between the two runs.
\label{fig:temp_comparison} }
\end{figure}

@@

%%%%%%%%%%%%
%%%%%%%%%%%
% Computational time-savings
%%%%%%%%%%%
%%%%%%%%%%%%
\newpage$\,$\newpage
\section{Computational Time-savings of AE}
\label{sec:speedups}

>>>>

Unfortunately, astrophysical convective systems are stiff in more than one 
timescale. Specifically, the Peclet number (Pe), the ratio of the thermal
diffusion timescale to the convective velocity timescale, is large.
In a high Pe system, many convective timescales must pass before 
the domain reaches thermal equilibrium, a steady state in which
evolution of the thermal structure of the convective region has ceased.
As the timestep size of implicit methods is bound to the fastest nonlinear flows,
(convection) fully implicit methods cannot be used to address this form of stiffness
\cite{viallet&all2011, viallet&all2013, viallet&all2016}.

Resolving dynamics in atmospheres which are sufficiently
thermally equilibrated therefore remains a challenging problem.
Solar convection is a prime example of this phenomenon, as
dynamical timescales in the solar convective zone are relatively short 
(convection overturns every $\sim 5$ minutes at the solar surface)
compared to the Sun's thermal evolution timescale, which is O($10^7$) years
\cite{stix2003}.  
In such a system, it is impossible to resolve the convective dynamics while also
meaningfully evolving the thermal structure of the system using
traditional timestepping techniques alone.
As modern simulations aim to model natural, high-Pe convection
in the high-Rayleigh-number (Ra) regime,
the thermal diffusion timescale ($t_{\kappa}$, defined in section \ref{sec:ae}) 
becomes intractably large compared to dynamical timescales 
such as the freefall time ($t_{\text{ff}}$, defined in section \ref{sec:experiment})
\cite{anders&brown2017}, 
\begin{equation}
\frac{t_{\kappa}}{t_{\text{ff}}} \propto (\text{Ra})^{1/2}.
\end{equation}
Furthermore, as dynamical and thermal timescales separate, 
simulations become more turbulent. Capturing appropriately resolved
turbulent motions requires finer grid meshes and smaller timesteps.
Thus, the progression of simulations into the high-Ra
regime of natural convection is slowed by two simultaneous effects: timestepping
through a single convective overturn time becomes more computationally expensive
and the number of overturn times required for systems to reach thermal equilibrium
grows.

@@

In this work, we study a method of achieving accelerated evolution of
convective simulations. We couple measurements of the dynamics of unequilibrated
convective simulations with knowledge about energy balances in the desired solution
to self-consistently adjust the mean vertical thermodynamic profile towards its evolved state. 
While a technique of this kind has been used previously \cite{hurlburt&all1986},
the details of implementation, the convergence properties, and whether or not the
accelerated evolution state corresponds to a standard evolution state are not documented.
In section \ref{sec:experiment}, we describe our convective simulations and
numerical methods. In section \ref{sec:ae}, we describe the procedure for 
achieving accelerated evolution. In
section \ref{sec:results}, we compare accelerated evolution solutions
to solutions obtained from standard evolution through a full thermal diffusion timescale. 
In section \ref{sec:speedups}, we compare the numerical cost of accelerated
solutions to standard solutions.
Finally,
in section \ref{sec:extensions}, we offer concluding remarks and
discuss extensions of the methods presented here.

@@

\section{Experiment}
\label{sec:experiment}
In this work we study a simple form of thermal convection:
incompressible \RB convection under the Oberbeck-Boussinesq approximation,
such that the fluid
has a constant kinematic viscosity ($\nu$), thermal diffusivity ($\kappa$), and coefficient
of thermal expansion ($\alpha$). The density of the fluid is a constant, $\rho_0$,
except in the buoyancy term, where it is $\rho = \rho_0(1  - \alpha T_1)$.
The gravitational acceleration, $\bm{g} = - g\bm{\hat{z}}$, is constant.
The equations of motion are \cite{spiegel&veronis1960}:
\begin{gather}
\DivU = 0, 
	\label{eqn:incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \bm{u}\cdot\grad\bm{u} =
-\frac{1}{\rho_0}\grad P - g( 1 - \alpha T_1)\hat{z} + \nu\grad^2\bm{u}, 
	\label{eqn:dim_bouss_momentum}
\\
\frac{\partial T_1}{\partial t} + \bm{u}\cdot\grad(T_0 + T_1) = \kappa\grad^2 T_1,
	\label{eqn:dim_bouss_energy}
\end{gather}
where $\bm{u} = u\bm{\hat{x}} + v\bm{\hat{y}} + w\bm{\hat{z}}$ is the velocity, 
$T = T_0(z) + T_1(x, y, z, t)$ are the initial and fluctuating components of temperature, 
and $P$ is the kinematic pressure. The initial temperature profile, $T_0$, decreases
linearly with height.
We non-dimensionalize these equations such that
length is in units of the layer height ($L_z$),
temperature is in units of the initial temperature jump across the layer ($\Delta T_0 = L_z \partial_z T_0$), 
and velocity is in units of the freefall velocity ($v_{\text{ff}} = \sqrt{\alpha g L_z^2 \partial_z T_0}$).
By these choices, one time unit is a freefall time ($t_{\text{ff}} = L_z/v_{\text{ff}}$).
We introduce a reduced kinematic pressure,
$\varpi \equiv (P / \rho_0 + \phi + |\bm{u}|^2 / 2) / v_{\text{ff}}^2$, where the gravitational
potential, $\phi$, is defined such that $\bm{g} = -\grad \phi$. 
In non-dimensional form, and substituting 
$\bm{u}\cdot\grad\bm{u} = \grad(|\bm{u}|^2/2) - \bm{u}\times(\grad\times\bm{u})$
and $\grad^2\bm{u} = -\grad\times(\grad\times\bm{u})$, Eqns.~\ref{eqn:dim_bouss_momentum} \& \ref{eqn:dim_bouss_energy}
become
\begin{align}
\frac{\partial \bm{u}}{\partial t} + \grad \varpi - T_1\hat{z} + \mathcal{R}\grad\times\bm{\omega} &= \bm{u}\times\bm{\omega},
	\label{eqn:bouss_momentum}
\\
\frac{\partial T_1}{\partial t} - \mathcal{P}\grad^2 T_1 + w \frac{\partial T_0}{\partial z} &= - \bm{u}\cdot\grad T_1,
	\label{eqn:bouss_energy}
\end{align}
where $\bm{\omega} = \grad \times \bm{u}$ is the vorticity.
The dimensionless control parameters $\mathcal{R}$ and $\mathcal{P}$ 
are set by the Rayleigh (Ra) and Prandtl (Pr) numbers,
\begin{equation}
\mathcal{R} \equiv \sqrt{\frac{\text{Pr}}{\text{Ra}}}, \qquad \mathcal{P} \equiv \frac{1}{\sqrt{\text{Pr}\,\text{Ra}}}, \qquad
\text{Ra} = \frac{g \alpha L_z^4 \partial_z T_0}{\nu\kappa} = \frac{(L_z\,v_{\text{ff}})^2}{\nu\kappa}, \qquad \text{Pr} = \frac{\nu}{\kappa}.
\end{equation}
We hold Pr $= 1$ constant throughout this work, such that $\mathcal{P} = \mathcal{R}$.
$\mathcal{P}$ and $\mathcal{R}$ are related to the inverse Reynolds and Peclet numbers of the
evolved flows.

In Eqns.~(\ref{eqn:incompressible}), (\ref{eqn:bouss_momentum}), \& (\ref{eqn:bouss_energy}),
linear terms are grouped on the left-hand side of the equations, while nonlinear terms
are found on the right-hand side. We timestep linear terms implicitly, and nonlinear
terms explicitly.
We utilize the 
Dedalus\footnote{\url{http://dedalus-project.org/}} 
pseudospectral framework \cite{burns&all2016} to evolve  
Eqns.~(\ref{eqn:incompressible}), (\ref{eqn:bouss_momentum}), \& (\ref{eqn:bouss_energy}) 
forward in time
using an implicit-explicit (IMEX), third-order, four-stage 
Runge-Kutta timestepping scheme RK443 \cite{ascher&all1997}. The code used to run the simulations
in this work is included in the supplemental materials as a zip file.

@@

The horizontally averaged profiles of the vertical conductive flux, 
F$_{\kappa} = \angles{-\kappa\grad(T_0 + T_1)}_{x,y}$, and the vertical convective enthalpy flux,
F$_{\text{E}} = \angles{w(T_0 + T_1)}_{x,y}$, are the basis of the AE method.
Here we use $\angles{}_{x,y}$ to represent a horizontal average. We measure
both of these fluxes early in a simulation, retrieving profiles such as
those shown in Fig.~\ref{fig:time_trace}b.
As the atmosphere evolves towards
the isothermal profile specified by the upper (cold) boundary condition, excess
temperature throughout the atmosphere must leave the domain. This excess thermal
energy leaves through the upper boundary, as seen in 
Fig.~\ref{fig:time_trace}b, where the amount of flux exiting at the top of the domain
is nearly 20 times larger the flux entering the bottom of the domain. Once the atmospheric
temperature profile reaches its evolved state, the flux entering the bottom boundary
is equal to the flux exiting through the upper boundary.  In general, this 
evolution is slow in SE (Fig.~\ref{fig:time_trace}a), but AE (Fig.~\ref{fig:time_trace}c)
can rapidly advance a system whose fluxes are in a strongly disequilibrium state (Fig.~\ref{fig:time_trace}b),
into a near-equilibrium state, as shown in Fig.~\ref{fig:time_trace}d. In
this final state both boundaries conduct the same amount of flux.
The converged fluxes achieved through AE are at most 5\% different from the SE solution, 
as shown in Fig.~\ref{fig:time_trace}e. 
This is a very small difference considering the short timescales on which convergence
is reached and the strongly disequilibrium state used to inform the AE process.

@@

To properly reduce the conductive flux, we examine the 
time-stationary state, horizontally- and time-averaged
Eqns.~(\ref{eqn:bouss_momentum}) and (\ref{eqn:bouss_energy}); after 
neglecting terms which vanish due to symmetry, these equations become
\vspace{-0.2cm}
\begin{gather}
\frac{\partial}{\partial z}\angles{\varpi}_{x,y} - \angles{T_1}_{x,y}\hat{z} = \angles{\bm{u}\times\bm{\omega}}_{x,y, \text{ ev}},
	\label{eqn:bouss_BVP_momentum}
\\
\frac{\partial}{\partial z}\text{F}_{\text{E, ev}} - \mathcal{P}\frac{\partial^2}{\partial z^2} \angles{T_1}_{x,y} = 0.
	\label{eqn:bouss_BVP_energy}
\end{gather}
Here, we construct  $\text{F}_{\text{E, ev}} = \xi \text{F}_{\text{E}}$ 
and $\angles{\bm{u}\times\bm{\omega}}_{x, y,\text{ ev}} =
\xi \angles{\bm{u}\times\bm{\omega}}_{x, y}$, where we scale 
$\angles{\bm{u}\times\bm{\omega}}_{x, y} \propto \bm{u}^2$, by $\xi$ because we
scaled $\bm{u}$ by $\sqrt{\xi}$ in order to reduce the convective enthalpy flux.
With these profiles, we solve for $\angles{\varpi}_{x,y}$ and $\angles{T_1}_{x,y}$.
Convective flows
are perturbations around a thermal profile defined by these equations in the evolved, 
statistically stationary state. Thus, the specification of
F$_{\text{conv, ev}}$ and $\angles{\bm{u}\times\bm{\omega}}_{x,y\text{, ev}}$
fully specifies the mean thermodynamic structure of the system. 
Our choice of fixing the temperature at the top boundary and the flux at the bottom
boundary ensures that there is a unique solution for $\angles{T_1}$ given the evolved
fluxes. If flux were fixed at both boundaries, there would be infinitely many
temperature solutions. If temperature were fixed at both boundaries, 
$\text{F}_{\text{B}}$ would not be precisely known \emph{a priori}.

Thus, the AE method occurs in the following steps. First, we pause a convective simulation and
construct $\xi(z,t)$ from measured flux profiles in the convective domain.
We solve a 1D boundary value problem (BVP) consisting of
Eqns.~(\ref{eqn:bouss_BVP_momentum}) \& (\ref{eqn:bouss_BVP_energy})
to obtain an evolved thermodynamic profile, and its corresponding conductive flux.
We multiply both the temperature perturbations around the mean and the
convective velocity flows by $\sqrt{\xi}$.
After adjusing the fields of a simulation in this manner, we continue timestepping forward.
For specifics on the precise implementation of the AE method, we refer
the reader to appendix \ref{appendix:recipe}.

@@

\begin{figure}[b]
\includegraphics[width=\textwidth]{./figs/nu_v_time.png}
\caption{The time evolution of Nu (black) and $\angles{T} - T_{\text{top}}$ (blue)
are shown for simulations at $S = 10^5$ in SE (a) and AE (b).
A moving average of Nu, using a centered boxcar with a span of 50 freefall time
units, is overplotted as a grey line. 
We see that the mean value of Nu reaches its equilibrium value quickly
compared to $\angles{T}$.
Regardless, fluctuations of Nu about the mean
value are much smaller in the thermally equilibrated state.
The times at which AE solves occur are denoted by arrows in (b),
and the grey boxed region is examined in more detail in Fig. \ref{fig:oscillating_plumes}.
\label{fig:nu_v_time} }
\end{figure}

@@

The Nusselt number (Nu) quantifies the efficiency of convective heat transport
and is defined as
\begin{equation}
\text{Nu} = \frac{\angles{F_{\text{conv}} + F_{\text{cond}}}}{\angles{F_{\text{cond, ref}}}}
 = \frac{\angles{wT - \mathcal{P}\partial_z T}}{\angles{- \mathcal{P} \partial_z T}},
\end{equation}
where the volume average of a quantity, $\eta$, is shown as $\angles{\eta}$.
The time evolution of Nu in AE and SE is compared to the mean temperature evolution
in Fig.~\ref{fig:nu_v_time}. In Fig.~\ref{fig:nu_v_time}a, we show the evolution
of the SE run at $S = 10^5$. The black trace shows the instantaneous value of
Nu, and the overplotted grey line shows a moving time average of Nu. The
time average is taken using a centered boxcar window whose width is 50 $t_{\text{ff}}$.
The mean value of Nu evolves towards its final value (dashed horizontal
grey line) rapidly compared to $\angles{T}-T_{\text{top}}$ (blue), 
which approaches its evolved state (dashed horizontal blue line) slowly.
As such, it is possible to measure the statistically stationary mean value of
Nu after only a few hundred simulated freefall times, 
as has been done previously \cite{stevens&all2010}. However, fluctuations in
Nu around the mean value at early times have both a larger magnitude and higher
frequency than the fluctuations in the equilibrated state.
In Fig.~\ref{fig:nu_v_time}b, we show the AE traces of Nu and
$\angles{T}-T_{\text{top}}$ at $S = 10^5$. The times at which AE solves occur are marked
by arrows. Nu reaches its evolved mean value slightly more rapidly than in the SE case,
and the frequency and magnitude of fluctuations in Nu away from the mean resemble the
final equilibrated state of SE.

When $S < 10^{3.67}$ in 2D and for all runs in 3D, 
the evolved system is characterized by a time-stationary value of Nu, and is thus
in a state of constant convective heat transport.
At larger $S$ in 2D, the value of Nu varies significantly over time even in the
equilibrated state (as seen in Fig.~\ref{fig:nu_v_time}). We examine the shaded
region of Fig.~\ref{fig:nu_v_time}b in more detail in the top left
panel of Fig.~\ref{fig:oscillating_plumes}, as well as a comparable time trace
at $S = 10^7$ (bottom, left panel). We find that these systems exhibit
large Nu during
states in which temperature fluctuations travel in their natural buoyant
direction (Fig.~\ref{fig:oscillating_plumes}, Ia \& IIa, where cold elements fall and hot elements rise).
However, when wrongly-signed temperature perturbations are entrained in an upflow or downflow
with oppositely signed fluid, Nu is suppressed (Fig.~\ref{fig:oscillating_plumes}, Ib and IIb, 
where warm fluid is pulled down by the downflow
lane, and cool fluid is drawn up by the upflow lane).
The plumes in these
systems naturally oscillate horizontally over time, switching between transport being dominated
by a counterclockwise cell, as pictured in Fig.~\ref{fig:oscillating_plumes}, and
a clockwise cell. Our choice of no-slip
boundary conditions prevents the fluid from entering a full domain shearing state 
\cite{goluskin&all2014}, and the
oscillatory nature of the plumes is stable. However, thanks to our choice of periodic
boundary conditions and despite the no-slip conditions, the full system of the
upflow and downflow plumes is free to slowly migrate to the left or right over time,
and we observe this phenomenon in our solutions.
The 2D SE simulations exhibit the same horizontally
oscillatory behavior as the AE solutions for the same initial conditions. 
This time-dependent behavior of Nu is not seen strongly in our 3D solutions,
however most 3D simulations we conducted were at low $S$ compared to the runs in
which this behavior was observed in 2D.

@@

\begin{figure}[t!]
\includegraphics[width=\textwidth]{./figs/temp_comparison.png}
\caption{Comparisons of the evolved thermodynamic states of an AE and SE run
at $S = 10^{5}$ are shown.  (a) Evolved horizontally- and time-averaged 
temperature profiles, as a function of height.
(b) Probability Distribution Functions (PDFs) and their integrated
Cumulative Distribution Functions (CDFs)
of point-by-point measurements of the temperature field.
(c) The percentage difference between the mean temperature profiles as a function of height.
The difference between the mean profiles is very small, O(0.5\%).
(d) $\KS{T}$, as defined in Eqn.~\ref{eqnks_profile}, is shown. The small
difference in the mean interior temperature between AE and SE
results in a large difference between the two temperature distributions near the values
of the temperature maxima.  The spread of temperature around the maxima, which includes the 
fluctuations that drive convection, are nearly identical between the two runs.
\label{fig:temp_comparison} }
\end{figure}

@@

%%%%%%%%%%%%
%%%%%%%%%%%
% Computational time-savings
%%%%%%%%%%%
%%%%%%%%%%%%
\section{Computational Time-savings of AE}
\label{sec:speedups}
